Holomorphic field theories and higher algebra

### 2024-09-16 draft.tex
continue to tinker with intro

\owen{mention some history (e.g., Manin! Donaldson-Thomas, Nekrasov) and some recent work (CPTVV, Pridham, etc) but then explain our focus: higher analogs of VOAs and finding mathematical analogs of physics}

@@

\brian{we need to make sure to cite nekrasov}

@@

The overarching goal of this paper is to introduce the reader to this emerging domain where physics and mathematics meet,
with an emphasis upon how key objects from physics, notably supersymmetric theories and conformal field theories, 
interface with complex geometry and especially with mathematical methods of rather recent vintage,
notably derived geometry and higher algebras.
A community is forming, involving both mathematicians and physicists,
to bring into sharp mathematical form some of the deep insights into gauge theory offered by physics.
There is a lot to explore here.

@@

For a systematic introduction to classical field theory in a style amenable to a complex geometer,
we suggest \cite{FreedDeligne},
although there are many other sources, such as \cite{}.
For a detailed development of field theory in dialogue with derived geometry,
see \cite{Steffens},
and for a motivational discussion, likely more accessible to physicists, see \cite{AlfonsoYoung?}.
In Section~\ref{sec: pert dfn} we offer a sharp definition well-suited to perturbative field theory 
(i.e., studying formal deformations of a fixed solution to the equations of motion),
which we will use to produce examples of higher algebras.

@@

\subsection{Perturbative field theories via derived deformations}
\label{sec: pert dfn}

@@

\begin{dfn}t 
A {\em holomorphic dg Lie algebra} on a complex manifold $X$ is
\begin{itemize}
\item a graded holomorphic vector bundle $V$ on $X$ whose sheaf of holomorphic functions is $\cV^{hol}$, and
\item a holomorphic differential operator $Q^{hol} \colon \cV^{hol} \to \cV^{hol}$ of cohomological degree one and a holomorphic bidifferential operator $[\cdot,\cdot] \colon \cV^{hol} \times \cV^{hol} \to \cV^{hol}$. 
\end{itemize}
This data is required to endow $(\cV^{hol}, Q^{hol}, [\cdot,\cdot])$ with the structure of a dg Lie algebra.
\end{dfn}

>>>>

\owen{mention some history (e.g., Manin, Donaldson-Thomas, Nekrasov) and some recent work (CPTVV, Pridham, Kapranov, etc) but then explain our focus: higher analogs of VOAs and finding mathematical analogs of physics}

@@

@@

\subsection{The aims of this survey}

Over the last few decades there has been a vigorous dialogue between complex geometry and theoretical physics,
around topics like mirror symmetry, enumerative algebraic geometry, moduli of Calabi-Yau manifolds, the geometric Langlands program, and integrable systems, among others.
Our focus will be narrow:
it is about higher dimensional versions of the relationship between geometry and algebraic structures like Lie algebras, Poisson algebras, and vertex algebras.
We will emphasize analogies to TFTs (such as the Gerstenhaber algebras appearing in mirror symmetry) and CFTs (such as the affine Lie algebras relating to moduli of $G$-bundles on Riemann surfaces),
but our primary interest is on complex manifolds above dimension one.
Experience suggests that higher dimensions require higher algebra,
i.e., algebraic methods that involve cochain complexes.
Thus differential graded (dg) algebras and categories,
as well as fellow travelers like derived geometry, appear.

A community has formed, involving both mathematicians and physicists, to explore and exploit these relationships,
and it continues to grow as there is a lot to explore here.
Examples of efforts in this direction include \cite{PTVV}, \cite{FHK}, \cite{Keviny things}, \cite{GaiottoEtAl}, \cite{Paquette}, \cite{Vicedo}, \cite{BeemBZEtAl}, 
where our own background and interests (and unfortunately, our ignorance) have shaped this list.

We will not discuss important directions like mirror symmetry, the geometric Langlands program, the BPS/CFT correspondence, Donaldson-Thomas theory, or other, more global and geometric, areas,
although we expect there are rich connections with the topics we do discuss.
Not including these topics is primarily because our core theme already swelled to more pages than we expected, 
because a concise discussion of any of those topics does not yet seem possible with respect to our theme,
and because of our lack of background in these areas.

@@

For a systematic introduction to classical field theory, 
in a style amenable to a complex geometer,
we suggest \cite{FreedDeligne}.
There are many other sources, such as \cite{}, because classical field theory is an old and deep subject.
Recently there is an effort to revisit field theory with insights from derived geometry.
See \cite{AlfonsoYoung?} for motivation and overview, 
but for a detailed development, see \cite{Steffens}.
In Section~\ref{sec: pert dfn} we offer a sharp definition well-suited to perturbative field theory 
(i.e., studying formal deformations of a fixed solution to the equations of motion),
which we will use to produce examples of higher algebras.

@@

\subsection{Perturbative field theory as derived deformation theory}
\label{sec: pert dfn}

@@

\begin{dfn} 
A {\em holomorphic dg Lie algebra} on a complex manifold $X$ is
\begin{itemize}
\item a graded holomorphic vector bundle $V$ on $X$ whose sheaf of holomorphic functions is $\cV^{hol}$, and
\item a holomorphic differential operator $Q^{hol} \colon \cV^{hol} \to \cV^{hol}$ of cohomological degree one and a holomorphic bidifferential operator $[\cdot,\cdot] \colon \cV^{hol} \times \cV^{hol} \to \cV^{hol}$. 
\end{itemize}
This data is required to endow $(\cV^{hol}, Q^{hol}, [\cdot,\cdot])$ with the structure of a dg Lie algebra.
\end{dfn}
